---
title: "Getting Started with didhelpers: Diagnosing Your DiD Design"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with didhelpers: Diagnosing Your DiD Design}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Overview

**didhelpers** provides lightweight diagnostic and visualization utilities for
staggered difference-in-differences (DiD) designs. It is not a new estimator ---
rather, it is the companion toolkit you wish you had when setting up a staggered
DiD analysis.

The package offers three core diagnostics and a convenience wrapper:
1. **Treatment timing plot** --- see when each unit gets treated.
2. **Pre-trends test** --- event-study regression to check the parallel trends assumption.
3. **Bacon decomposition summary** --- understand which 2x2 comparisons drive your TWFE estimate.
4. **Full diagnostics report** --- run everything at once.

## Setup

```{r setup}
library(didhelpers)
```

We will use the package's built-in simulated dataset, which has 30 units
observed over 15 time periods with staggered treatment adoption across four
cohorts (including a never-treated group). The true ATT is 2.0, and there are
no pre-trends by construction.

```{r data}
data("staggered_data")
head(staggered_data)
table(staggered_data$cohort)
```

## 1. Plot Treatment Timing

The first step in any staggered DiD analysis is understanding the treatment
variation in your panel. `plot_treatment_timing()` produces a heatmap showing
when each unit switches from untreated to treated:

```{r treatment-timing}
plot_treatment_timing(staggered_data, unit, time, treat)
```

Units are grouped by treatment cohort on the y-axis. This makes it easy to
spot the staggered adoption pattern and identify never-treated units.

## 2. Test Pre-Trends

The parallel trends assumption is central to any DiD design.
`test_pretrends()` runs an event-study regression with unit and time fixed
effects and returns pre-treatment lead coefficients:
```{r pretrends}
pt <- test_pretrends(staggered_data, unit, time, outcome, treat, n_lags = 3)
pt$coefficients
```

The joint F-test checks whether all pre-treatment coefficients are jointly
zero:

```{r ftest}
pt$f_test
```

Since our simulated data has no true pre-trends, the lead coefficients should
be small and statistically insignificant. The `autoplot()` method produces the
classic event-study coefficient plot:

```{r es-plot}
ggplot2::autoplot(pt)
```

## 3. Bacon Decomposition

Goodman-Bacon (2021) showed that the two-way fixed effects (TWFE) estimator
is a weighted average of all possible 2x2 DD comparisons. Some of these
comparisons may use already-treated units as controls, leading to potential
bias. `bacon_weight_summary()` provides a clean summary:

```{r bacon}
bs <- bacon_weight_summary(staggered_data, unit, time, outcome, treat)
bs$summary
```

The scatter plot shows each 2x2 comparison's estimate against its weight,
colored by comparison type:

```{r bacon-plot}
bs$plot
```

## 4. Full Diagnostics Report

For a quick overview, `did_diagnostics_report()` runs all three diagnostics
at once:

```{r report}
report <- did_diagnostics_report(staggered_data, unit, time, outcome, treat)
report
```

Individual components are accessible via `report$treatment_plot`,
`report$pretrends`, and `report$bacon`.

## 5. Printing and Summaries

All didhelpers objects have clean `print()` and `summary()` methods modelled
after the output style of **fixest**:

```{r print-pt}
pt
```

```{r print-bs}
bs
```

The `summary()` method provides additional detail (R-squared, HHI of weights):

```{r summary-pt}
summary(pt)
```

```{r summary-bs}
summary(bs)
```

## 6. Broom Integration

The `tidy()`, `glance()`, and `augment()` methods enable seamless integration
with **modelsummary** and tidyverse workflows:

```{r broom-tidy}
tidy(pt)
glance(pt)
```

```{r broom-bacon}
tidy(bs, type = "summary")
glance(bs)
```

See `vignette("publishing-output")` for full details on table export and
figure customization.

## Relationship to Other Packages

**didhelpers** is a *diagnostic companion*, not a replacement for any
estimator package. Here is how it relates to the ecosystem:

| Package | Relationship | Notes |
|:--------|:-------------|:------|
| **bacondecomp** | Hard dependency | `bacon_weight_summary()` wraps `bacondecomp::bacon()`, adding S3 class, tidy output, broom methods, and gt/kable tables. |
| **fixest** | Hard dependency | `test_pretrends()` uses `fixest::feols()` internally. fixest's own `iplot()` is more flexible for fixest-native workflows; didhelpers bundles the event-study + F-test + broom integration into a single call. |
| **HonestDiD** | Complementary | Provides sensitivity analysis for parallel trends violations. No overlap -- could be used *after* `test_pretrends()` flags concerns. |
| **did2s** | No overlap | Implements the Gardner (2022) imputation estimator. didhelpers diagnoses TWFE; did2s replaces it. |
| **DIDmultiplegt** | No overlap | Suite of robust DiD estimators. didhelpers could diagnose data *before* running these estimators. |
| **staggered** | Minimal overlap | Provides Callaway-Sant'Anna and Sun-Abraham estimators with their own plot methods. These plot *estimates* from robust estimators; didhelpers plots *TWFE diagnostics*. |

## Next Steps
These diagnostics help you decide whether a standard TWFE estimator is
appropriate for your data. If the Bacon decomposition reveals substantial
weight on problematic comparisons (e.g., already-treated as controls), or if
pre-trends tests fail, consider robust estimators such as those in the
[DIDmultiplegt](https://github.com/chaisemartinPackages/did_multiplegt_dyn)
family of packages.
