---
title: "Publishing Tables and Figures from didhelpers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Publishing Tables and Figures from didhelpers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Overview

This vignette demonstrates how to turn **didhelpers** diagnostic output into
publication-quality tables and figures suitable for academic papers, slide decks,
and reports.

We cover four output paths:

1. **gt / kableExtra tables** via `as_table()`
2. **broom-style tidy output** via `tidy()`, `glance()`, `augment()`
3. **Regression tables** via `modelsummary` (LaTeX, Word, HTML)
4. **Customized ggplot2 figures** for publication

## Setup

```{r setup}
library(didhelpers)
data("staggered_data")

# Run diagnostics
pt <- test_pretrends(staggered_data, unit, time, outcome, treat, n_lags = 3)
bs <- bacon_weight_summary(staggered_data, unit, time, outcome, treat)
```

## 1. Publication Tables with gt

The `as_table()` function produces formatted tables for direct inclusion in
RMarkdown and Quarto documents.

```{r gt-pretrends, eval = requireNamespace("gt", quietly = TRUE)}
as_table(pt, format = "gt")
```

```{r gt-bacon, eval = requireNamespace("gt", quietly = TRUE)}
as_table(bs, format = "gt")
```

For kableExtra output (better for LaTeX/PDF documents), use `format = "kable"`:

```{r kable-pretrends, eval = requireNamespace("kableExtra", quietly = TRUE)}
as_table(pt, format = "kable")
```

## 2. Broom Integration

The `tidy()`, `glance()`, and `augment()` methods make didhelpers objects work
seamlessly in tidyverse workflows.

### Tidy coefficients

```{r tidy}
tidy(pt)
```

### Model-level statistics

```{r glance}
glance(pt)
```

### Fitted values and residuals

```{r augment}
head(augment(pt))
```
### Bacon decomposition

```{r tidy-bacon}
# Per-type summary
tidy(bs, type = "summary")

# All 2x2 comparisons
tidy(bs, type = "two_by_twos")

# Decomposition-level stats
glance(bs)
```

## 3. LaTeX and Word Export via modelsummary

With broom methods in place, `modelsummary` works automatically. This is the
recommended path for regression tables in academic papers.

```{r modelsummary-default, eval = requireNamespace("modelsummary", quietly = TRUE)}
# Interactive HTML/gt display
export_pretrends_table(pt)
```

```{r modelsummary-latex, eval = requireNamespace("modelsummary", quietly = TRUE)}
# LaTeX output (paste into your .tex file)
export_pretrends_table(pt, output = "latex")
```

To export directly to a Word document:

```r
# Save to .docx (requires flextable)
export_pretrends_table(pt, output = "my_table.docx")
```

### Why modelsummary?

| Package      | S3 extensible | LaTeX | Word | Maintained |
|:-------------|:-------------:|:-----:|:----:|:----------:|
| modelsummary | Yes           | Yes   | Yes  | Yes        |
| stargazer    | No            | Yes   | No   | No (2022)  |
| huxtable     | Partial       | Yes   | Yes  | Yes        |

**modelsummary** is the best fit because it uses `tidy()` and `glance()`
automatically, supports all output formats, and integrates with gt and
kableExtra.

## 4. Publication-Ready Figures

All plot methods return ggplot2 objects that can be further customized.

### Event-study plot

```{r pub-esplot}
library(ggplot2)
autoplot(pt) +
  theme_classic(base_size = 12) +
  labs(title = NULL, caption = NULL) +
  theme(legend.position = "none")
```

### Bacon decomposition plot

```{r pub-bacon}
autoplot(bs) +
  theme_classic(base_size = 12) +
  labs(title = NULL) +
  theme(legend.position = "bottom")
```

### Composite report figure

```{r pub-composite, eval = requireNamespace("patchwork", quietly = TRUE), fig.width = 10, fig.height = 8}
report <- did_diagnostics_report(staggered_data, unit, time, outcome, treat)
plot(report)
```

## Summary

| Task                        | Function / Method                          |
|:----------------------------|:-------------------------------------------|
| Console summary             | `print()`, `summary()`                     |
| Tidy coefficient table      | `tidy(pt)`                                 |
| Model-level stats           | `glance(pt)`                               |
| gt / kableExtra table       | `as_table(pt)`, `as_table(bs)`             |
| LaTeX regression table      | `export_pretrends_table(pt, output = "latex")` |
| Word regression table       | `export_pretrends_table(pt, output = "file.docx")` |
| modelsummary (direct)       | `modelsummary::modelsummary(list(pt))`     |
| Publication ggplot2 figures | `autoplot(pt) + theme_classic()`           |
| Composite diagnostics plot  | `plot(report)` (requires patchwork)        |
