---
title: "didhelpers Demonstration"
author: "Stefano Graziosi"
date: "`r Sys.Date()`"
format:
  html:
    theme: cosmo
    toc: true
    code-fold: true
execute:
  warning: false
  message: false
---

## Introduction

This document provides a comprehensive demonstration of the **`didhelpers`** R package. Designed as a lightweight toolkit for staggered Difference-in-Differences (DiD) designs, it allows researchers to easily diagnose, visualize, and summarize their empirical strategies.

Below, we load the package, leverage the built-in dataset, and sequentially apply all four core diagnostic functions.

## Setup & Data

First, let's load the required packages and the sample dataset.

```{r setup}
# Load the newly installed package
library(didhelpers)
library(ggplot2)

# Load the built-in staggered dataset
data("staggered_data")

# Preview the data structure
head(staggered_data)
```

## Function 1: `plot_treatment_timing()` 🗺️

Before jumping into regressions, we visualize when treatment adoption occurs across different cohort groups.

```{r treatment-timing, fig.width=8, fig.height=5}
# Generates a heatmap of treatment rollout
plot_treatment_timing(staggered_data, unit, time, treat)
```

## Function 2: `test_pretrends()` 📈

To visually inspect the parallel trends assumption, we run an event-study regression with time and unit fixed effects. This function calculates the coefficients and automatically generates a pre-trend F-test.

```{r pretrends, fig.width=8, fig.height=5}
# Run the event study
pt_model <- test_pretrends(staggered_data, unit, time, outcome, treat)

# Inspect raw coefficients and standard errors
print(pt_model$coefficients)

# Render the coefficient plot
autoplot(pt_model) +
  labs(title = "Event Study: Assessing Parallel Trends",
       subtitle = "Reference: Period before treatment (t = -1)")
```

## Function 3: `bacon_weight_summary()` ⚖️

When running Two-Way Fixed Effects (TWFE) with staggered timing, the final estimator is a weighted average of many underneath 2x2 comparisons. We use the Goodman-Bacon decomposition to summarize exactly how much weight is placed on 'clean' comparisons versus potentially biased 'already-treated' comparisons.

```{r bacon-weights, fig.width=8, fig.height=5}
# Run the Bacon Decomposition
bacon_decomp <- bacon_weight_summary(staggered_data, unit, time, outcome, treat)

# View the tabular summary of weights by comparison type
knitr::kable(bacon_decomp$summary, digits = 3)

# Visualize the underlying estimates vs. their assigned weight
bacon_decomp$plot +
  labs(title = "Goodman-Bacon Decomposition Weights")
```

## Function 4: `did_diagnostics_report()` 📑

Finally, `didhelpers` provides a single wrapper function to generate a consolidated report, compiling the results of all the functions above simultaneously.

```{r report}
# Run the all-in-one report
final_report <- did_diagnostics_report(staggered_data, unit, time, outcome, treat)

# Print the final report structure
print(final_report)
```

## Conclusion

The `didhelpers` package provides a standardized, easily reportable toolkit to diagnose modern DiD estimators. This Quarto template perfectly showcases its ease of use under a tidy framework!
